# Dates and times

Load libraries.

```{r, warning=FALSE, message=FALSE}
library(dplyr)      # for data manipulation
library(lubridate)  # for date and time functions
```

Here are some useful date and time functions. When using strptime, you will have to specify the date and time components. For help on the formats available, run ?strptime.

```{r}

# get current system date
Sys.Date() 

# get current system datetime, with the time zone
Sys.time()

# convert a character datetime to POSIXct
# specify the components of the datetime when using strptime

x <- "2024/11/15 12:00:00"
( z <- as.POSIXct(strptime(x, "%Y/%m/%d %H:%M:%S")) ) 

class(z) # note the formats of z

```

Adding date and time variables to a data frame is easy using lubridate functions. When calculating time differences, it is best to use the difftime function rather than straight subtraction.

```{r}
# first, let's make an example data frame (test) with a date variable formats

today <- as.POSIXct(Sys.time())

test <- data.frame(date=seq(today,today+5*60,60)) 

class(test$date) # note the format

glimpse(test) # glimpse the data frame as a tibble


# now let's use lubridate functions to extract date & time components

test1 <- test %>%
  mutate(year = year(date),
         yday = yday(date),
         month = month(date),
         day = day(date),
         hour=hour(date),
         min=minute(date),
         sec=second(date),
         tomorrow=date+days(1)) %>%
  mutate(difftime = difftime(tomorrow,date,units="hours"),
         difftime_h = as.numeric(difftime))

test1

```

