# Sun event times

Load libraries

```{r, warning=FALSE, message=FALSE}

library(dplyr)      # for data frame manipulation
library(suntools)   # for sunriset, crepuscule
library(lubridate)  # for date and time functions
library(data.table) # for setDT

```

Functions in suntools provide everthing you need to get sunrise, sunset, and other sun event times. One caveat is that these functions do not seem to perform well in dplyr. Therefore, we will 1) create a new variable using the assignment operator and 2) add the data table format to our data frame and then perform manipulations on the data table. 

It's useful to look at an example before attempting these functions on your own data frame.

```{r}

# see example for sunriset

example(sunriset)

```

Now let's try suntools functions on a data frame. Be very careful with your time zones! 

```{r}

# make example data frame (df) with dates, lat, and lon
( today <- Sys.Date() ) # get today's date in UTC

# create a data frame with dates, and convert time to tz
df <- data.frame(date=as.POSIXct(seq(today,today+10,by=1),tz="EST"),
                 lat=42,
                 lon=-80)

# put lat and lon into a matrix
coords <- matrix(c(df$lon,df$lat),nrow=nrow(df))

# get sunrise in tz of df$date
# here we are using the assignment operator, not dplyr functions
df$sunrise <- sunriset(
  coords,
  df$date,
  direction = c("sunrise"),
  POSIXct.out = TRUE
)$time

# confirm that the tz is correct
df$sunrise[1:6]


# get sunset in tz of df$date
df$sunset <- sunriset(
  coords,
  df$date,
  direction = c("sunset"),
  POSIXct.out = TRUE
)$time
df$sunset[1:6]

```

As an alternative, we can add a data table format to the data frame and then use data table (dt) syntax for data manipulations. dt syntax works faster on large data sets.


```{r}

# add dt format to the data frame & check
setDT(df)
class(df)

# sunset in tz of df$date
df[,sunrise1 := sunriset(matrix(c(lon, lat),
                                        ncol = 2, byrow = FALSE), date,
                                 POSIXct.out=T,
                                 direction='sunrise')[2]][]
# sunset in tz of df$date
df[,sunset1 := sunriset(matrix(c(lon, lat),
                                ncol = 2, byrow = FALSE), date,
                         POSIXct.out=T, direction='sunset')[2]][]

# confirm these are correct
df$sunrise1[1:6]
df$sunset1[1:6]

```

Now let's get crepuscular times:

- civil twilight: sun is 6 degrees below horizon
- nautical twilight: sun is 12 degrees below horizon
- astonomical twilight: sun is 18 degrees below horizon

```{r}

# astronomical twilight at dawn

df$dawn = crepuscule(
  coords,
  df$date,
  solarDep=18,
  direction = c("dawn"),
  POSIXct.out = TRUE
)$time

# astronomical twilight at dusk
df$dusk = crepuscule(
  coords,
  df$date,
  solarDep=18,
  direction = c("dusk"),
  POSIXct.out = TRUE
)$time

df$dawn[1:6] # confirm tz
df$dusk[1:6]

# keep variables of interest and view
df1 <- df %>%
  select(date,lat,lon,sunrise,sunset,dawn,dusk)

head(df1)

```


