# Maps

This chapter shows how to create a basic map of a study area in the Laurentian Great Lakes region. The map will also include the Great Lakes shoreline from a shapefile I had previously acquired (and now hard to find).

Load libraries

```{r, warning=FALSE, message=FALSE}
library(ggmap)
library(ggplot2)
library(sf)  # needed for st_transform
```

The first step is to get a stadiamap API key (see <https://docs.stadiamaps.com/authentication/)>. My recommendation is to save this key in an R script call stadiakey.R with the following code:

```{r, eval=FALSE}

# stadiamaps API key
ggmap::register_stadiamaps(key = "YOUR KEY")

```

Then, source your script to register your API key in your work session. Once this is done you can acquire and save a base map.

```{r, eval=FALSE}

source("PATHNAME/stadiakey.R") # runs the script

# get a map

long_point <- get_stadiamap(bbox=c(left=-83.5,
                                   right=-79,
                                   bottom=41.5,
                                   top=45.5),
                            maptype = "stamen_terrain_background",
                            zoom = 8,
                            color = "color")

ggmap(long_point) # view before saving

save(long_point,file="./data/long_point.rda")

```


Now you're ready to gather the features.

```{r}

load("./data/long_point.rda") # load previously saved base layer


# get the Great Lakes shoreline layer
greatLakes <- read_sf(dsn="./data/", 
                      layer="gl_mainlakes_wgs84_rp")

# transform into correct crs (coordinte reference system)
greatLakes_4326 <- st_transform(greatLakes, 4326) 

```


Make the map. This is pretty basic and there is much more you can do!


```{r, warning=FALSE, message=FALSE}

map <- ggmap(long_point)+
  geom_sf(data = greatLakes_4326,
          mapping = aes(geometry = geometry),
          color = "darkolivegreen",
          fill = "#98B2CD",
          inherit.aes = FALSE)+
  ylab("Latitude")+
  scale_x_continuous(name="Longitude",
                     limits=c(-83.5,-79),
                     expand=c(0,0))+
  coord_sf(xlim = c(-83.5, -79), 
           ylim = c(41.5, 45.5), expand = TRUE) +
  theme_bw()+
  theme(legend.position = "none")+
  theme(axis.text=element_text(size=8),
        text = element_text(size =8),
        axis.title=element_text(size=10),
        strip.text = element_text(size = 12))+
  annotate("text", label = "Lake Erie", fontface="italic",
    x = -81.15, y = 42.2, size = 3)

map

```